MSE Scenarios for short and long lived

### 2025-04-15 Scenarios.Rmd
trying to add data poor scenario, breaking

@@

Project at F40%

@@

info <- generate_basic_info(n_stocks = 1, 
                            n_regions = 1, 
                            n_indices = 2, # spring and fall
                            n_fleets = 1, 
                            n_seasons = 1, # don't need seasons
                            base.years = year_start:year_end,
                            n_feedback_years = MSE_years,
                            life_history = "long",
                            n_ages = 20,
                            # We want a bit higher F in the historical period
                            F_info = list(F.year1 = 0.1, Fhist = "F-H-L", Fmax = 2.5, Fmin = 1, change_time = 0.5, user_F = NULL),
                            catch_info = list(catch_cv = 0.1, catch_Neff = 100),
                            index_info = list(index_cv = 0.1, index_Neff = 100, fracyr_indices = c(0.295, 0.812), q = 0.2),
                            fracyr_spawn = 0.625)

@@

@@

mods[[1]] = loop_through_fn(om = om_with_data,
                            em_info = info, 
                            random = random,
                            M_em = M, # use OM M
                            sel_em = sel, # use OM sel
                            NAA_re_em = NAA_re_em, # use rec assumed random around the mean instead, help runtime (est B-H is difficult)
                            move_em = NULL,
                            age_comp_em = "multinomial",
                            # Here is the correct code: separate.em = FALSE also works for one-area model
                            em.opt = list(separate.em = FALSE, separate.em.type = 1, 
                                          do.move = FALSE, est.move = FALSE),
                            assess_years = assess.years, 
                            assess_interval = assess.interval, 
                            base_years = base.years,
                            year.use = 20, # number of years of data you want to use in the assessment model
                            add.years = TRUE, # extends assessment time series instead of moving window of year.use years
                            seed = 123,
                            save.sdrep = FALSE, 
                            save.last.em = TRUE,
                            FXSPR_init = 0.01) # IMPORTANT!

@@

@@

>>>>

Looks like ~0.2 is reasonable for redfish in ICELAND, higher here.

Say its decent ~0.3

@@

Project at F40% (75% F40% is HCR 1)

@@

info <- generate_basic_info(n_stocks = 1, 
                            n_regions = 1, 
                            n_indices = 2, # spring and fall
                            n_fleets = 1, 
                            n_seasons = 1, # don't need seasons
                            base.years = year_start:year_end,
                            n_feedback_years = MSE_years,
                            life_history = "long",
                            n_ages = 20,
                            # We want a bit higher F in the historical period
                            F_info = list(F.year1 = 0.1, Fhist = "F-H-L", Fmax = 2.5, Fmin = 1, change_time = 0.5, user_F = NULL),
                            catch_info = list(catch_cv = 0.1, catch_Neff = 100),
                            index_info = list(index_cv = 0.3, index_Neff = 100, fracyr_indices = c(0.295, 0.812), q = 0.2),
                            fracyr_spawn = 0.625)

@@

# -------------------
# IMPORTANT!#
# -------------------
# Change initial numbers at age in the input for OM
# ini.NAA <- matrix(NA, n_ages, n_stocks)
# ini.NAA[,1] <- c(20:1) # Use actually initial numbers at age from assessment model?
# 
# input$par$log_N1[] = 0
# for (i in 1:n_regions) {
#   input$par$log_N1[i,i,] = log(ini.NAA[,i])
# }

@@

mods[[1]] = loop_through_fn(om = om_with_data,
                            em_info = info, 
                            random = random,
                            M_em = M, # use OM M
                            sel_em = sel, # use OM sel
                            NAA_re_em = NAA_re_em, # use rec assumed random around the mean instead, help runtime (est B-H is difficult)
                            move_em = NULL,
                            age_comp_em = "multinomial",
                            # Here is the correct code: separate.em = FALSE also works for one-area model
                            em.opt = list(separate.em = FALSE, separate.em.type = 1, 
                                          do.move = FALSE, est.move = FALSE),
                            assess_years = assess.years, 
                            assess_interval = assess.interval, 
                            base_years = base.years,
                            year.use = 30, # number of years of data you want to use in the assessment model
                            add.years = TRUE, # extends assessment time series instead of moving window of year.use years
                            seed = 123,
                            save.sdrep = FALSE, 
                            save.last.em = TRUE,
                            FXSPR_init = 0.01) # IMPORTANT!

@@

```{r}
agg_index_sigma = input$data$agg_index_sigma
agg_index_sigma[31:60,] = 0.6 # Increase CV for both survey indices in the feedback period
index_Neff = input$data$index_Neff
index_Neff[31:60,] = 50 # Decrease ESS for both survey indices in the feedback period

#alternate years fall and spring surveys
remove_agg = TRUE # remove a aggregate index for some years
remove_agg_pointer = c(1,2) # both
remove_agg_years = matrix(data=c(seq(31,60,2), seq(32,60,2)), nrow=15, ncol=2) #alternating years by survey
remove_paa = TRUE # Also remove age comp for that index 
remove_paa_pointer = c(1,2) # both
remove_paa_years = matrix(data=c(seq(31,60,2), seq(32,60,2)), nrow=15, ncol=2) #alternating years by survey

input <- update_input_index_info(input, agg_index_sigma, index_Neff,
                                 remove_agg, remove_agg_pointer, remove_agg_years,
                                 remove_paa, remove_paa_pointer, remove_paa_years) # Update input file

agg_catch_sigma = input$data$agg_catch_sigma
agg_catch_sigma[31:60,] = 0.2 #double catch CV in the feedback period
catch_Neff = input$data$catch_Neff
catch_Neff[31:60] = 50

input <- update_input_catch_info(input, agg_catch_sigma, catch_Neff)

Generate new om and dataset, keep same assessment interval as above

```{r}
om <- fit_wham(input, do.fit = F, do.brps = T, MakeADFun.silent = TRUE)
om_with_data <- update_om_fn(om, seed = 123, random = random)
```

Test new EM

mods[[2]] = loop_through_fn(om = om_with_data,
                            em_info = info, 
                            random = random,
                            M_em = M, # use OM M
                            sel_em = sel, # use OM sel
                            NAA_re_em = NAA_re_em, # use rec assumed random around the mean instead, help runtime (est B-H is difficult)
                            move_em = NULL,
                            age_comp_em = "multinomial",
                            # Here is the correct code: separate.em = FALSE also works for one-area model
                            em.opt = list(separate.em = FALSE, separate.em.type = 1, 
                                          do.move = FALSE, est.move = FALSE),
                            assess_years = assess.years, 
                            assess_interval = assess.interval, 
                            base_years = base.years,
                            year.use = 30, # number of years of data you want to use in the assessment model
                            add.years = TRUE, # extends assessment time series instead of moving window of year.use years
                            seed = 123,
                            save.sdrep = FALSE, 
                            save.last.em = TRUE,
                            FXSPR_init = 0.01) # IMPORTANT!

Visualize 2 mods (modified from https://lichengxue.github.io/whamMSE/03.Management-Strategy-Evaluation.html#12_Compare_model_performance)

This compares the base case,good surveys in the future (1) with degraded data, alternating year seasonal surveys with lower quality sampling (catch too) in the future (2)

par(mfrow = c(1,2))
SSB_s1 <- lapply(mods, function(mod) mod$om$rep$SSB[,1])
plot(SSB_s1[[1]], type = "l", col = "blue", ylab = "SSB", xlab = "Year", ylim=c(0,100000), main = "OM SSB") 
colors <- c("red","green","purple","orange")
for (i in length(SSB_s1)) {
  lines(SSB_s1[[i]], col = colors[i-1],lty = i)
}
legend("topleft",legend = paste0("EM ", 1:2), col = c("blue",colors),lty=1:5, cex=0.8)

Catch_s1 <- lapply(mods, function(mod) mod$om$rep$pred_catch[,1])
plot(Catch_s1[[1]], type = "l", col = "blue", ylab = "Catch", xlab = "Year", ylim=c(0,30000), main = "OM Catch")
colors <- c("red","green","purple","orange")
for (i in length(Catch_s1)) {
  lines(Catch_s1[[i]], col = colors[i-1],lty = i)
}

@@

Do this later...

For now, do
